Slitasje og kjørespor

### 2023-02-02 slitasje.Rmd

---
title: "Slitasje og kjørespor"
output:
  html_document
editor_options: 
  markdown: 
    wrap: 72
---

@@

## Introduction

@@

3.  Covert [ANO points to polygons](#ANO-points-to-poly)

4.  For each polygon, extract the NiN-variable that has the lowest value
    (worst-rules rule), e.g. like Fig. \@ref(fig:slitasje-naturetype).

@@

10. Sample 10 00 random points from an ecosystem delineation and extract
    indicator values and sd at those points.

@@

The indicator uses a data set from a standardised field survey of nature
types. You can read more about the preliminary analyses
[here](#naturtype). See also the [official
site](https://www.miljodirektoratet.no/ansvarsomrader/overvaking-arealplanlegging/naturkartlegging/naturtyper/)
of the Environmental Agency. I also import a data set called
[ANO](#ANO-import), which you can read about
[here](https://www.miljodirektoratet.no/ansvarsomrader/overvaking-arealplanlegging/miljoovervaking/overvakingsprogrammer/natur-pa-land/arealrepresentativ-naturovervakning-ano/)

@@

The nature type mapping is not random and cannot be said to be area
representative. The ANO dataset however, is area representative. The
data is from 2018 to present. The data from one field season usually
becomes available early the following year.

@@

The variables are recorded on a coarse four-sped ordinal scale (Fig.
\@ref(fig:four-step)).

@@

The indicator is tagged to the *Økologisk egenskap* called
**Primærproduksjon** (Primary productivity). This is tentative, and
perhaps *abiotiske forhold* is more suited. But the thought behind the
chiose is that *sliatsje* affect the potential for primary productivity.

@@

The indicator is tagged as a **Physical state characteristics**
indicator. This is quite clear. It's a tpe of abiotic characteritic that
has to do with the physical structure rather than the chemical
composition.

@@

Slitasje is not thought to exhibit collinearity with any other indicator at
the present.

@@

The upper reference value is 0 (no *slitasje*), and the lower reference
value is 100%. Note that 100% slitasje does not mean that all the area
must be scared or damaged, but that all hypothetical quadrats around a
point is affected (the variable is frequency-driven, and not
amount-driven). The threshold for good ecosystem condition is 10%
damage. This is on the conservative side. Perhaps is should be set
somewhat higher, like 20 or 30%.

@@

Uncertaintiesare estimated for aggregated indicator values by
bootstrapping individual indicator values 1000 times and calculating a
distribution of area weighted means. When aggregating a second time,
from homogeneous impact areas to accounting areas, we assume a normal
distribution around the indicator values, with the already mentioned
errors, and sample n times from these and combine the resamples into an
a new, area weighted, distribution. The errors for the accounting areas
thus represents both the spatial variation, and the precision, of the
indicator values within the accounting areas.

>>>>

# (PART\*) INDICATORS{-}

@@

## Introduction
This indicator represent human cause terrain damage from hiking or off-road vehicle traffic. The data come from a combination non-random field surveys and area-representative surveys. Therefore extra care is taken not to over-extrapolate, and we use local reference area in the form of _homogeneous impact areas_.

@@

3.  Convert [ANO points to polygons](#ANO-points-to-poly)

4.  For each polygon, extract the NiN-variable that has the lowest value
    (one-out all-out principle), e.g. Fig. \@ref(fig:slitasje-naturetype).

@@

10. Sample 1000 random points from an ecosystem delineation and extract
    indicator values and sd at those points.

@@

The indicator uses a data set from a standardised field survey of nature
types. You can read more about the preliminary analyses
[here](#naturtype). See also the [official
site](https://www.miljodirektoratet.no/ansvarsomrader/overvaking-arealplanlegging/naturkartlegging/naturtyper/)
of the Environment Agency. I also import a data set called
[ANO](#ANO-import), which you can read about
[here](https://www.miljodirektoratet.no/ansvarsomrader/overvaking-arealplanlegging/miljoovervaking/overvakingsprogrammer/natur-pa-land/arealrepresentativ-naturovervakning-ano/)

@@

The nature type mapping is not random and cannot be said to be area
representative. The ANO data set however, is area representative. The
data is from 2018 to present. The data from one field season usually
becomes available early the following year.

@@

The variables are recorded on a coarse four-step ordinal scale (Fig.
\@ref(fig:four-step)).

@@

The indicator is tagged to the *Økologisk egenskap* called
**Primærproduksjon** (Primary productivity). This is tentative, and
perhaps *abiotiske forhold* is more suited. But the thought behind the
choise is that *sliatsje* affect the potential for primary productivity.

@@

The indicator is tagged as a **Physical state characteristics**
indicator. This is quite clear. It's a type of abiotic characteristic that
has to do with the physical structure rather than the chemical
composition.

@@

_Slitasje_ is not thought to exhibit collinearity with any other indicator at
the present.

@@

* Upper = 0%

* Threshold = 10%

* Lower = 100%

The upper reference value is 0 (no *slitasje*), and the lower reference
value is 100%. Note that 100% _slitasje_ does not mean that all the area
must be scarred or damaged, but that all hypothetical quadrats around a
point is affected (the variable is frequency-driven, and not
amount-driven). The threshold for good ecosystem condition is 10%
damage. This is on the conservative side. Perhaps is should be set
somewhat higher, like 20 or 30%.

@@

Uncertainties/errors are estimated for aggregated indicator values by
bootstrapping individual indicator values 1000 times and calculating a
distribution of area weighted means. When aggregating a second time,
from homogeneous impact areas to accounting areas, we assume a normal
distribution around the indicator values, with the already mentioned
errors, and sample _n_ times from these and combine the resamples into an
a new, area weighted, distribution. The errors for the accounting areas
thus represents both the spatial variation, and the precision, of the
indicator values within the accounting areas.
